Exploring Income Inequality Dynamics in Canada: A Geospatial and Temporal Analysis

### 2024-04-06 paper/paper.qmd
upload

---
title: "Exploring Income Inequality Dynamics in Canada: A Geospatial and Temporal Analysis"
subtitle: "Insights from 1976 to 2021"
author: 
  - Sirui Tan
thanks: "Code and data are available at: https://github.com/siru1366/income.git."
date: today
date-format: long
abstract: "This paper investigates income inequality in Canada from 1976 to 2021 using data from various surveys. Through exploratory data analysis and regression modeling, we uncover increasing income inequality trends over time and variations across different provinces. Our findings reveal nuanced relationships between temporal trends and regional disparities, and income distribution dynamics. Through rigorous analysis, we uncover two key findings: a consistent rise in income inequality over time, significant regional disparities across provinces. Understanding these dynamics is crucial for policymakers to develop targeted strategies aimed at mitigating income inequality and promoting social and economic equity in Canada."
format: pdf
toc: true
number-sections: true
bibliography: references.bib
---

@@

# Introduction
Income inequality is a pressing socioeconomic issue that has garnered significant attention worldwide, with Canada being no exception. Over the past few decades, Canada has witnessed a widening gap in per capita after-tax income, raising concerns about economic disparities and social justice. Understanding the drivers behind this trend is crucial for policymakers and researchers alike, as it provides insights into the factors influencing income distribution and informs strategies to mitigate inequality.

This paper aims to investigate the intricate dynamics of income inequality within Canada, focusing on the period from 1976 to 2021. Leveraging a comprehensive dataset spanning multiple surveys, including the Survey of Consumer Finances (SCF), the Survey of Labour and Income Dynamics (SLID), and the Canadian Income Survey (CIS), we embark on a rigorous exploration of income trends across different provinces and over time. By examining key variables such as year, geographical location, income decile, and income range, we seek to uncover patterns and nuances in income distribution dynamics.

@@

Data cleaning and analysis were conducted using the open-source statistical programming language `R` [@citeR], leveraging functionalities from the `tidyverse` [@tidyverse], `ggplot2` [@ggplot2], `dplyr` [@dplyr], `readr` [@readr], `tibble` [@tibble],`stringr`  [@stringr], `haven` [@haven], `janitor` [@janitor], `knitr` [@knitr].

@@

2.  **Geographical location**: The dataset includes data from all provinces in Canada, providing a comprehensive view of income distribution dynamics across the entire country. This geographic diversity enables the examination of regional variations and disparities in income levels and inequality measures.

@@

We develop two separate models to delve into the dynamics of income inequality and range variations in Canada.

@@

Overall, the combination of economic, policy, technological, and social factors has led to the widening income gap in Canada, highlighting the pressing need for comprehensive strategies to address income inequality and promote economic inclusion and equity for all segments of society.

@@

Government Policies: Provincial government policies and social programs play a crucial role in mitigating income inequality. PEI's government may implement policies focused on social welfare, affordable housing, and income support programs to address economic disparities and ensure a more equitable distribution of wealth @peiGovernmentPolicy. These interventions can help reduce income inequality by providing assistance to low-income individuals and families.

@@

Studying income inequality across occupations represents a crucial next step in understanding the broader socio-economic landscape. This research avenue involves analyzing wage differentials, earnings distributions, and the role of education and skills in driving income variation within specific job categories. Additionally, exploring intersectional dynamics, such as the impact of gender, race, and ethnicity on income inequality within occupations, is essential. Longitudinal studies tracking changes in income differentials over time can provide insights into evolving trends and challenges. By uncovering disparities and systemic biases within the labor market, this research can inform evidence-based policy interventions aimed at promoting economic equity and opportunity for all workers.

>>>>

---
title: "Exploring Income Inequality Dynamics in Canada: A Geospatial and Temporal Analysis"
subtitle: "Insights from 1976 to 2021"
author: 
  - Sirui Tan
thanks: "Code and data are available at: https://github.com/siru1366/income.git."
date: today
date-format: long
abstract: "This paper investigates income inequality in Canada from 1976 to 2021 using data from various surveys. Through exploratory data analysis and regression modeling, we uncover increasing income inequality trends over time and variations across different provinces. Our findings reveal nuanced relationships between temporal trends and regional disparities, and income distribution dynamics. Through rigorous analysis, we uncover two key findings: a consistent rise in income inequality over time, significant regional disparities across provinces. Understanding these dynamics is important for policymakers to develop targeted strategies aimed at mitigating income inequality and promoting social and economic equity in Canada."
format: pdf
toc: true
number-sections: true
bibliography: references.bib
---

@@

# Introduction
Income inequality is a notable socioeconomic issue that has gained attention worldwide, including in Canada. Over recent decades, Canada has seen a widening gap in per capita after-tax income, raising concerns about economic disparities and social justice. Understanding the drivers behind this trend is important for policymakers and researchers, providing insights into the factors influencing income distribution and informing strategies to mitigate inequality.

This paper aims to investigate the intricate dynamics of income inequality within Canada, focusing on the period from 1976 to 2021. Using a wide-ranging dataset from surveys like the Survey of Consumer Finances (SCF), the Survey of Labour and Income Dynamics (SLID), and the Canadian Income Survey (CIS), we thoroughly investigate income trends across provinces and over time. By examining key variables such as year, geographical location, income decile, and income range, we seek to uncover patterns and nuances in income distribution dynamics.

@@

Data cleaning and analysis were conducted using the open-source statistical programming language `R` [@citeR], utilizing functionalities from the `tidyverse` [@tidyverse], `ggplot2` [@ggplot2], `dplyr` [@dplyr], `readr` [@readr], `tibble` [@tibble],`stringr`  [@stringr], `haven` [@haven], `janitor` [@janitor], `knitr` [@knitr].

@@

2.  **Geographical location**: The dataset covers all provinces in Canada, giving a complete view of income distribution dynamics across the country. This broad geographic coverage enables the examination of regional variations and disparities in income levels and inequality measures.

@@

We develop two separate models to find the dynamics of income inequality and range variations in Canada.

@@

In summary, a range of economic, policy, technological, and social factors have contributed to the widening income gap in Canada. This highlights the urgent need for effective strategies to address income inequality and promote economic inclusion and equity for all segments of society.

@@

Government Policies: Provincial government policies and social programs play a vital role in mitigating income inequality. PEI's government may implement policies focused on social welfare, affordable housing, and income support programs to address economic disparities and ensure a more equitable distribution of wealth @peiGovernmentPolicy. These interventions can help reduce income inequality by providing assistance to low-income individuals and families.

@@

Studying income inequality across occupations represents a important next step in understanding the broader socio-economic landscape. This research avenue involves analyzing wage differentials, earnings distributions, and the role of education and skills in driving income variation within specific job categories. Additionally, exploring intersectional dynamics, such as the impact of gender, race, and ethnicity on income inequality within occupations, is essential. Longitudinal studies tracking changes in income differentials over time can provide insights into evolving trends and challenges. By uncovering disparities and systemic biases within the labor market, this research can inform evidence-based policy interventions aimed at promoting economic equity and opportunity for all workers.
